---
title: "Final Analysis of Impact Sound Level Data"
author: "Michael Parzinger"
date: "`r Sys.Date()`"
output: 
  html_document:
    number_sections: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)


recalculate <- FALSE

realDataAvailable <- TRUE

library(SoundModeling) # https://github.com/Michi-Parz/SoundModeling
if (realDataAvailable)
  library(SoundModelingData)
library(rstan)
library(stringr)

library(PupillometryR)
library(ggridges)

data("ntp_data")

ntp_share <- ntp_data
```


# Informations for own use


Please note: If *recalculate* is set to TRUE, it will take longer, but this is necessary for the first pass!


*realDataAvailable* TRUE is only possible if the R package *SoundModelingData* has been installed, which contains the actual data. However, this package is not publicly available!




# ISO 12354-2 Model

Assumptions:

$\rho = 2300$ kg/$m^3$, $t = 0.2$ m, $f_0=80$ Hz, $f_\mathrm{ref}=1000$ Hz, $q=35$, $\eta_\mathrm{int}=0.01$, C=$0.5$, $c_\mathrm{L}=3500$ m/s, und volume $V= 50$

Also, here is the dynamic maximum radiation level, because it makes the curve look more natural. (Different ceiling, so this is acceptable)


```{r}
testValuesFull <- impact_sound_standard(thickness = 0.2,
                                       reduction_factor = 35,
                                      resonance_freq = 80,
                                      volume = 50,
                                      freq = freq,
                                      max_rl_type = "dynamic")

lnCompare <- impact_sound_raw(thickness = 0.2, density = 2300, volume=50,
                         offset = 155, freq = freq)


lnTestDF <- data.frame("fpHz" = freq,
             "IMso" = testValuesFull, "Ln" = lnCompare)
```

```{r}
func_1b <- function(f, b0,b1, b2, bp){
  b0+b1*log10(f) + b2 * pmax(log10(f/bp), 0)
}


NLMsegTheo <- nlsLM(IMso ~ func_1b(fpHz, b0, b1, b2, bp),
      data = lnTestDF, start = c(b0=53 ,b1 =1.6 ,b2 = 35, bp =80))

summary(NLMsegTheo)


segRegCoeff <- coefficients(NLMsegTheo)
```

```{r}
gg_basic(
  lnTestDF,
  aes(y = IMso, x = fpHz, color = "L'n Model"),
  ymin = 0, ymax = 70,
  isl = T, lab = F
)+geom_line() + geom_point()+
  geom_line(aes(y = func_1b(fpHz,
                            segRegCoeff[1],
                            segRegCoeff[2],
                            segRegCoeff[3],
                            segRegCoeff[4]),
                color = "Regression"))+
  labs(color = "")+theme(
    legend.position = c(0.4, 0.2),
    legend.justification = c(0.5, 0.5),
    legend.background = element_rect(fill = "transparent"),
    legend.spacing.y = unit(-0.7, "cm"),
    panel.grid.major = element_line(linewidth = 1))+
  geom_line(aes(y = lnCompare, color = "Ln Model"))



```



```{r}
gg_basic(
  lnTestDF,
  aes(y = IMso, x = fpHz, color = "L'n Model"),
  ymin = 0, ymax = 70,
  isl = T, lab = F
)+geom_line() + geom_point()+
  geom_line(aes(y = func_1b(fpHz,
                            segRegCoeff[1],
                            segRegCoeff[2],
                            segRegCoeff[3],
                            segRegCoeff[4]),
                color = "Regression"))+
  labs(color = "")+theme(
    legend.position = c(0.4, 0.2),
    legend.justification = c(0.5, 0.5),
    legend.background = element_rect(fill = "transparent"),
    legend.spacing.y = unit(-0.7, "cm"),
    panel.grid.major = element_line(linewidth = 1))
```



# Data Adjustment

```{r}

ntp_data$ID <- 1:nrow(ntp_data)

ntp_data <- ntp_data[!ntp_data$seltsam_mbbm,]

ntp_original <- ntp_data

lin_adjust <- apply(
  ntp_data[,13:33],
  1, linear_continuation, orientation = 3
)


ntp_o_long <- tidyr::pivot_longer(
  ntp_data,
  cols = 13:33,
  names_to = "fpHz",
  values_to = "NTP"
)


#ntp_between <- ntp_data
#ntp_between[,13:33] <- (ntp_data[,13:33] + t(lin_adjust))/2

ntp_data[,13:33] <- t(lin_adjust)


# ntp_between <- tidyr::pivot_longer(
#   ntp_between,
#   cols = 13:33,
#   names_to = "fpHz",
#   values_to = "NTP"
# )


ntp_long <- tidyr::pivot_longer(
  ntp_data,
  cols = 13:33,
  names_to = "fpHz",
  values_to = "NTP"
)

ntp_o_long <- ntp_o_long[
  !ntp_o_long$seltsam_mbbm,
  c("ID", "fpHz", "NTP", "E_VolA")
]


# ntp_between <- ntp_between[
#   !ntp_long$seltsam_mbbm,
#   c("ID", "fpHz", "NTP", "E_VolA")
# ]

ntp_long <- ntp_long[
  !ntp_long$seltsam_mbbm,
  c("ID", "fpHz", "NTP", "E_VolA")
]






ntp_long$fpHz <- ntp_o_long$fpHz <- rep(freq, 142)


ntp_between <- ntp_long
ntp_between$NTP <- (ntp_long$NTP + ntp_o_long$NTP)/2


```

```{r}
gg_o <- gg_basic(ntp_o_long, aes(x = fpHz, y = NTP, group = ID),
         ymin = -20, ymax = 70, isl = T, lab = F)+
  geom_line(alpha = 0.5)+
  labs(title = "Original")

gg_a <- gg_basic(ntp_long, aes(x = fpHz, y = NTP, group = ID),
         ymin = -20, ymax = 70, isl = T, lab = F)+
  geom_line(alpha = 0.5)+
  labs(title = "Adjusted")


gg_bonus <- gg_basic(ntp_between, aes(x = fpHz, y = NTP, group = ID),
         ymin = -20, ymax = 70, isl = T, lab = F)+
  geom_line(alpha = 0.5)+
  labs(title = "Mean")


ggarrange(gg_o, gg_a)

ggarrange(gg_o, gg_a, gg_bonus, ncol = 3)

```

# Effect of Data Fitting on the Breakpoint

The breakpoint model is run repeatedly using the original data and a dataset up to and including 2500 Hz. The question then is how much the parameters differ from those of the fitted dataset.

```{r}
apply(ntp_original[,13:33]<=10, 2, sum)
```

Why 2500 Hz? Up to 1600 Hz, all values are greater than 10 dB. At 2000 Hz, only two measurements are affected, and at 2500 Hz, only 15 measurements are affected. At 3150 Hz, however, there are already 57 measurements, which is over $40\ \%$.

```{r}
func_1b <- function(f, b0,b1, b2, bp){
  b0+b1*log10(f) + b2 * pmax(log10(f/bp), 0)
}

func_2b <- function(f, b0,b1, b2, b3, bp, bp_diff){
  bp2 <- bp + bp_diff
  b0+b1*log10(f) + b2 * pmax(log10(f/bp), 0) + b3 * pmax(log10(f/bp2), 0)
}

```

```{r}

starvals <- coefficients(NLMsegTheo)

mod_adata_1b <- nlsLMcov(NTP~func_1b(fpHz, b0,b1,b2, bp), ntp_long,
                   start = starvals,
             cov_type = "tvAR",
             maxlag = 1 
             )

mod_odata_1b <- nlsLMcov(NTP~func_1b(fpHz, b0,b1,b2, bp), ntp_o_long,
                   start = starvals,
             cov_type = "tvAR",
             maxlag = 1 
             )

mod_mdata_1b <- nlsLMcov(NTP~func_1b(fpHz, b0,b1,b2, bp), ntp_between,
                   start = starvals,
             cov_type = "tvAR",
             maxlag = 1 
             )


mod_rdata_1b <- nlsLMcov(NTP~func_1b(fpHz, b0,b1,b2, bp),
                         ntp_o_long[ntp_o_long$fpHz <= 2500,],
                   start = starvals,
             cov_type = "tvAR",
             maxlag = 1, curve_len = 18
             )



data_influence_list <- list(
  "Lin. continuation" = mod_adata_1b,
  "Original" = mod_odata_1b,
  "Reduced" = mod_rdata_1b,
  "Averaged" = mod_mdata_1b
)



vgl_data <- data.frame(
  "fpHz" = c(freq,freq,
             #freq,
             freq[1:18]),
  "Fit" = c(mod_adata_1b$fitted[1:21],
    mod_odata_1b$fitted[1:21],
    #mod_mdata_1b$fitted[1:21],
    mod_rdata_1b$fitted[1:18]),
  "Data" = c(
    rep(c("Lin. continuation", "Original"
          #,"Average"
          ), each = 21),
    rep("Reduced", 18)
  )
)

gg_basic(vgl_data, aes(x = fpHz, y = Fit, color = Data, group = Data),
         ymin = 0, ymax = 60, isl = T)+
  geom_line()+
  labs(color = "Data basis")+
  theme(
    legend.position = c(0.4, 0.2),
    legend.justification = c(0.5, 0.5),
    legend.background = element_rect(fill = "transparent"),
    legend.spacing.y = unit(-0.7, "cm"),
    panel.grid.major = element_line(linewidth = 1))


```

```{r}

lapply(data_influence_list, summary)

sapply(data_influence_list, coefficients)
```


# LOO comparison segmented regression

```{r}

loolist <- readRDS("Saves/LOOlist_Fix")


loo::loo_compare(loolist)

loo_df <- data.frame(loo::loo_compare(loolist))[,1:2]

rnames <- row.names(loo_df)

loo_df <- cbind("K" = stringr::str_sub(rnames, 2,2),
                "L" = stringr::str_sub(rnames, 4,4),
                  loo_df)

row.names(loo_df) <- NULL

knitr::kable(loo_df, digits = c(0,0,1,1))

```

The TVAR(3) model with three segments is the most suitable.


# Model with two breakpoints


$$
 y(f) = \beta_0 + \beta_1 \lg(f) + \beta_{2} \max(\lg(f/\tau_1),0) + \beta_{3} \max(\lg(f/\tau_2),0)
$$




## Priors



-   $\beta|\Sigma,X_\tau\ \sim \mathcal{N}_4(b, (X_\tau\Sigma^{-1}X_\tau)^{-1})$

-   $b=(38.6,4.88,-18.8,-18.8)^T$

-   $X_\tau$: Design matrix so that $E[y_i|\beta,\tau] = X_\tau\beta$

-   $\delta_1 \sim \mathrm{Exp}(1/0.36)$

-   $\delta_2 \sim \mathrm{Exp}(1/914)$

-   $\tau_1 = 50 + \delta_1$

-   $\tau_2 = \tau_1+\delta_2 $



```{r}

bp1 <- 86
bp2 <- 1000


x1 <- rep(1, 21)
x2 <- log10(freq)
x3 <- ifelse(freq > bp1, 1,0) * log10(freq/bp1)
x4 <- ifelse(freq > bp2, 1,0) * log10(freq/bp2)

X <- cbind(x1,x2,x3,x4)




```





## Diagnosis




```{r}
model3seg <- readRDS("Saves/ImSo_Fix_2breaks_tvAR_3")

fit3seg <- rstan::extract(model3seg)




set.seed(260326)
y_rep <- array(dim = c(4000,142, 21))

for (i in 1:4000) {
  y_rep[i,,] <- mvtnorm::rmvnorm(142, fit3seg$mu[i,], fit3seg$cov_mat[i,,])
}

fit3seg$y_rep <- y_rep

```



```{r}
# n_eff
summary(head(summary(model3seg)$summary[,"n_eff"], 6+57+21))
# Rhat
summary(head(summary(model3seg)$summary[,"Rhat"], 6+57+21))


check_hmc_diagnostics(model3seg)

```





## Model curves




```{r}

ntp_fit <- data.frame(
    "ID" = -1,
    "fpHz" = freq,
    "NTP" = apply(fit3seg$mu, 2, mean),
    "Type" = "Model: K=3"
  )





```






```{r}
    # geom_line(
    #     mapping = aes(x = fpHz, y = NTP),
    #     data = ntp_mean, size = 0.8
    # )+

gg_basic(ntp_long,
         ymin = -20, ymax = 70,
         isl = T, lab = F)+
    geom_line(aes(x = fpHz, y = NTP, group = ID), alpha = 0.1)+
  geom_line(
        mapping = aes(x = fpHz, y = NTP),
        data = ntp_fit, linewidth = 0.8, color = "red"
    )


```



## Residuals




```{r}


exp_white_res <- data.frame(
  "fpHz" = rep(freq,142),
  "Res" = c(t(apply(fit3seg$st_resid, 2:3, mean))),
  "Fit" = rep(apply(fit3seg$mu,2,mean), 142)
)


resfreq <- ggWhiteRes(exp_white_res, aes(x = as.factor(fpHz), y = Res),
                      txtsize = 3, tikz = T)+
  labs(x = expression(paste("Frequency ",
                                                        italic(" f"), " / Hz")))+
  theme(axis.text.x = element_text(angle =40, hjust = 1))
    # scale_x_discrete( minor_breaks = c(freq,6300,8000,10000),
    #                   breaks = c(63,125,250,500,1000,2000,4000,8000),
    #                   labels = c(63,125,250,500,1000,2000,4000,8000),
    # )

resfreq


```



Notable observation


```{r}

outlier <- which.min(exp_white_res$Res)
outlierID <- ntp_long[outlier,]$ID


ntp_long$outlier <- ifelse(ntp_long$ID == outlierID, T, F)


whiteResOutlier <- gg_basic(ntp_long, aes(x = fpHz, y = NTP, group = ID, color = outlier),
         ymin = -20, ymax = 70, isl = T, lab = F)+
  geom_line(aes(alpha = outlier))+
  geom_point(aes(shape = outlier))+
  scale_color_manual(values = c("darkgrey", "black"))+
  scale_alpha_manual(values = c(0.25,1))+
  scale_shape_manual(values = c(NA,16))+
  theme(legend.position = "none")#+
  #geom_vline(xintercept = c(400,500,630), linetype = "dotted")


whiteResOutlier

```


## Residuals Chisq

```{r}



chi_curve <- apply(fit3seg$st_resid, 1:2, \(x){
  sum(x^2)
})

chi_freq <- apply(fit3seg$st_resid, c(1,3), \(x){
  sum(x^2)
})

chiWallQuantiles <- apply(chi_curve, 2,
                         quantile,
                         probs = c(0.5, 0.95, 0.99,1))

chiFreqQuantiles <- apply(chi_freq, 2,
                         quantile,
                         probs = c(0.5, 0.95, 0.99,1))


```

```{r}

chiCurveDf <- data.frame(
  "Wall" = rep(1:142, each = 4),
  "ChiSQ" = c(chiWallQuantiles)
)
chiFreqDf <- data.frame(
  "Freq" = rep(freq, each = 4),
  "ChiSQ" = c(chiFreqQuantiles)
)


```


```{r}
ggplot(chiCurveDf)+
  aes(x = Wall, y = ChiSQ)+
  geom_point()+
  geom_hline(
     yintercept = qchisq(c(0.5,0.95,0.99),21),
     color = "blue"
  )+
  geom_hline(
     yintercept = max(rchisq(4000, 21)),
     color = "lightblue"
  )


gg_basic(chiFreqDf,ymin = 0, ymax = 200, sound = F)+
  aes(x = Freq, y = ChiSQ)+
  geom_point()+
  geom_hline(
     yintercept = qchisq(c(0.5,0.95,0.99),142),
     color = "blue"
  )+
  geom_hline(
     yintercept = max(rchisq(4000, 142)),
     color = "lightblue"
  )

```





## Betas


Note: In the calculations in STAN, $\ln$ was always used instead of $\lg$!
Therefore, the corresponding coefficients here must be multiplied by the factor $\ln(10)$

```{r}

fix_betas <- data.frame(fit3seg$fe_betas)
names(fix_betas) <- paste("b", 0:3, sep = "")

fix_betas[,2:4] <- log(10)*fix_betas[,2:4]


gg_kernel <- ggplot(fix_betas)


# b anstatt bquote(Intercept~beta[0]), weil tikz

ggb0 <- gg_kernel+
  aes(y=b0, x = "")+
  geom_flat_violin() +
  geom_boxplot()+
  labs(y = "dB", x = "Intercept b0")
  
ggb1 <- gg_kernel+
  aes(y=b1, x = "")+
  geom_flat_violin() +
  geom_boxplot()+
  labs(y = "dB", x = "Slope b1")

ggb2 <- gg_kernel+
  aes(y=b2, x = "")+
  geom_flat_violin() +
  geom_boxplot()+
  labs(y = "dB", x = "Slope change b2")

ggb3 <- gg_kernel+
  aes(y=b3, x = "")+
  geom_flat_violin() +
  geom_boxplot()+
  labs(y = "dB", x = "Slope change b3")

ggarrange(ggb0,ggb1, ggb2, ggb3)


```

```{r}

fix_tau <- data.frame(fit3seg$b_points)

names(fix_tau) <- c("tau1", "tau2")




gg_kernel <- ggplot(fix_tau)


# b instead bquote(Intercept~beta[0]), becaus of tikz

ggt1 <- gg_kernel+
  aes(y=tau1, x = "")+
  geom_flat_violin() +
  geom_boxplot()+
  labs(y = "Hz", x = "Breakpoint t1")
  
ggt2 <- gg_kernel+
  aes(y=tau2, x = "")+
  geom_flat_violin() +
  geom_boxplot()+
  labs(y = "Hz", x = "Breakpoint t2")



ggarrange(ggt1,ggt2)


```





## Bedingte Standardabweichung

```{r}
quantiles <- c(0.025, 0.5, 0.975)
```


```{r}


# fit3seg$unconditional_st_dev also exists


stdev_bayes <- fit3seg$st_dev

stdev_b_quant <- apply(stdev_bayes, 2, quantile, probs = quantiles)

stdev_b_df <- data.frame(
  t(stdev_b_quant),
  fpHz = freq,
  Method = "Bayesian"
)
names(stdev_b_df)[1:3] <- paste("q", c(2.5,50,97.5), sep = "")


# Combined
stdev_df <- stdev_b_df
```


```{r}

ggplot(stdev_df)+
  aes(x = fpHz)+
  geom_line(aes(y = q50))+
  geom_ribbon(aes(ymin = q2.5, ymax = q97.5), alpha = 1/4)+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  labs(y = "Conditional standard deviation",
       x = "Frequency f / Hz")+
  theme(legend.position = "bottom")


```

## Marginal standard deviation

```{r}
quantiles <- c(0.025, 0.5, 0.975)
```


```{r}


# fit3seg$unconditional_st_dev also exists


stdev_bayes <- fit3seg$unconditional_st_dev

stdev_b_quant <- apply(stdev_bayes, 2, quantile, probs = quantiles)

stdev_b_df <- data.frame(
  t(stdev_b_quant),
  fpHz = freq,
  Method = "Bayesian"
)
names(stdev_b_df)[1:3] <- paste("q", c(2.5,50,97.5), sep = "")


# Combined
stdev_df <- stdev_b_df
```


```{r}

ggplot(stdev_df)+
  aes(x = fpHz)+
  geom_line(aes(y = q50))+
  geom_ribbon(aes(ymin = q2.5, ymax = q97.5), alpha = 1/4)+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  labs(y = "Marginal standard deviation / dB",
       x = "Frequency f / Hz")+
  theme(legend.position = "bottom")


```

## AR Coefficients

Hmm, how should I best describe them... there are 57 of them, after all...


```{r}


# Bayes
arcoeff_bayes <- fit3seg$shift_coeff


b_quant <- apply(arcoeff_bayes, 2, quantile, probs = quantiles)

ar_b_df <- data.frame(
  t(b_quant),
  fpHz = c(freq[-1],freq[-c(1:2)],freq[-c(1:3)]),
  Lag = c(rep(1,20), rep(2,19), rep(3,18))
)
names(ar_b_df)[1:3] <- paste("q", quantiles*100, sep = "")


# Combined
ar_df <- ar_b_df
```


```{r}
# 
ggplot(ar_df)+
  aes(x = fpHz, group = Lag,
      fill = as.factor(Lag),
      color = as.factor(Lag))+
  geom_line(aes(y = q50))+
  geom_ribbon(aes(ymin = q2.5, ymax = q97.5), alpha = 1/4)+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  labs(y = "AR coefficient",
       x = "Frequency f / Hz",
       fill = "Lag", color = "Lag")+
  theme(legend.position = "bottom")


```

## Correlation


```{r}
mean_corrmat <- apply(fit3seg$corr_mat, 2:3, mean)

row.names(mean_corrmat) <- colnames(mean_corrmat) <- paste(freq, "Hz")


corrplot::corrplot(mean_corrmat, "color", tl.col = "black")

```




## Sample qq



```{r}

chiWall <- apply(fit3seg$st_resid, 1:2, \(x){sum(x^2)})
chiFreq <- apply(fit3seg$st_resid, c(1,3), \(x){sum(x^2)})



```


```{r}

MCMCchiWall <- MCMCqqPlotsChi(chiWall, 21, 100)
MCMCchiFreq <- MCMCqqPlotsChi(chiFreq, 142, 100)

```





```{r}



ggplot(MCMCchiWall)+
  aes(x= TheoQuantiles,group = Index)+
  geom_line(aes(y= EmpQuantiles), alpha = 1/3, col = "lightblue")+
  geom_line(aes(y = TheoQuantiles))+
  geom_line(aes(y = TheoQuantiles + 2*StDevQuantile), linetype = "dotted")+
  geom_line(aes(y = TheoQuantiles - 2*StDevQuantile), linetype = "dotted")


ggplot(MCMCchiFreq)+
  aes(x= TheoQuantiles,group = Index)+
  geom_line(aes(y= EmpQuantiles), alpha = 1/3, col = "lightblue")+
  geom_line(aes(y = TheoQuantiles))+
  geom_line(aes(y = TheoQuantiles + 2*StDevQuantile), linetype = "dotted")+
  geom_line(aes(y = TheoQuantiles - 2*StDevQuantile), linetype = "dotted")


```

## PPC One-Third Ocatave Band


```{r}

alphas <- c(0.5,0.05,0.01)

yRepQ <- apply(fit3seg$y_rep, 2:3, quantile,
               probs = c(alphas/2, 1-alphas/2))

yRepQdf <- data.frame(t(data.frame(yRepQ)))

names(yRepQdf) <- c(
  paste("L", 1:length(alphas), sep = ""),
  paste("U", 1:length(alphas), sep = "")
)


ppcDF <- cbind(
  ntp_long[order(ntp_long$fpHz),], yRepQdf
)


```



```{r}
# gg_basic(ppcDF, aes(x = fpHz, group = ID, y = NTP),
#          ymin = -30, ymax = 80)+
#   geom_line(aes(color = "Data"))+
#   geom_ribbon(aes(ymin = L3, ymax = U3, color = "99 p"),
#               alpha = 0.01, fill = NA)+
#     geom_ribbon(aes(ymin = L2, ymax = U2, color = "95 p"),
#                 alpha = 0.01, fill = NA)+
#     geom_ribbon(aes(ymin = L1, ymax = U1, color = "50 p"),
#                 alpha = 0.01, fill = NA)+
#   scale_color_manual(values = c("green", "blue", "orange", "black"))+
#   #scale_fill_manual(values = c("lightgreen","lightblue","grey"))+
#     labs(color = "", fill = "PPC region", y = "Observation")+
#     theme(legend.position = "bottom")
```

```{r}
ppcFreq <- gg_basic(ppcDF, aes(x = fpHz, group = ID, y = NTP),
         ymin = -30, ymax = 80, isl = TRUE, lab = TRUE)+
  geom_line(aes(color = "Data"))+
  geom_line(aes(y = L3, color = "99 p"), alpha = 0.1)+
  geom_line(aes(y = U3, color = "99 p"), alpha = 0.1)+
  geom_line(aes(y = L2, color = "95 p"), alpha = 0.1)+
  geom_line(aes(y = U2, color = "95 p"), alpha = 0.1)+
  geom_line(aes(y = L1, color = "50 p"), alpha = 0.1)+
  geom_line(aes(y = U1, color = "50 p"), alpha = 0.1)+
  scale_color_manual(
    values = c("green", "blue", "orange", "black"),
    guide = guide_legend(override.aes = list(alpha = 1)))+
  #scale_fill_manual(values = c("lightgreen","lightblue","grey"))+
    labs(color = "Predictive interval")
    #theme(legend.position = "bottom")

ppcFreq

```



```{r}
mat1 <- OutOfRange(ppcDF, "NTP", "L1", "U1", result = "matrix")
mat2 <- OutOfRange(ppcDF, "NTP", "L2", "U2", result = "matrix")
mat3 <- OutOfRange(ppcDF, "NTP", "L3", "U3", result = "matrix")


expected <- 142*alphas


oorBias1 <- apply(mat1, 2, sum) - expected[1]
oorBias2 <- apply(mat2, 2, sum) - expected[2]
oorBias3 <- apply(mat3, 2, sum) - expected[3]

oorBias1sd <- oorBias1/sqrt(142*alphas[1]*(1-alphas[1]))
oorBias2sd <- oorBias2/sqrt(142*alphas[2]*(1-alphas[2]))
oorBias3sd <- oorBias3/sqrt(142*alphas[3]*(1-alphas[3]))


prfct <- function(ex, diff, prop){
   (1-pbinom(ex+diff,142,prop)+ pbinom(ex-diff,142,prop))
}


prob1 <- prfct(expected[1], diff = abs(oorBias1), alphas[1])
prob2 <- prfct(expected[2], diff = abs(oorBias2), alphas[2])
prob3 <- prfct(expected[3], diff = abs(oorBias3), alphas[3])



oorDf <- data.frame(
  "fpHz" = rep(freq, 3),
  "OORbias" = c(oorBias1,oorBias2,oorBias3),
  "OORbiasSd" = c(oorBias1sd,oorBias2sd,oorBias3sd),
  "alpha" = paste(rep(1-alphas, each = 21), "p"),
  "probs" = c(prob1,prob2,prob3)
)





ggplot(oorDf)+
  aes(x = fpHz, y = OORbias, color = alpha)+
  geom_point()+
  geom_line()+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  scale_color_manual(
    values = c("green", "blue", "orange"),
    breaks = c("0.5 p", "0.95 p", "0.99 p"))+
  labs(y = "OOR - Expected",
       x = expression(paste("Frequency ",
                             italic(" f"), " / Hz")),
       color = "Predictive interval")+
  theme(legend.position = "bottom")


ppcFreqSt <- ggplot(oorDf)+
  aes(x = fpHz, y = OORbiasSd, color = alpha)+
  geom_point()+
  geom_line()+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  scale_y_continuous(breaks = -3:3)+
  scale_color_manual(
    values = c("green", "blue", "orange"),
    breaks = c("0.5 p", "0.95 p", "0.99 p"))+
  labs(y = "(OOR - Expected) / Standard deviation",
       x = expression(paste("Frequency ",
                             italic(" f"), " / Hz")),
       color = "Predictive interval  ")+
  theme(legend.position = "bottom")+
  coord_cartesian(ylim = c(-3.1,3.1))

ppcFreqSt

ppcFreqBinom <- ggplot(oorDf)+
  aes(x = fpHz, y = probs, color = alpha)+
  geom_point()+
  geom_line(linetype = "dotted")+
  scale_x_continuous(trans = log2_trans(),
                       minor_breaks = c(freq,6300,8000,10000),
                       breaks = c(63,125,250,500,1000,2000,4000,8000),
                       labels = c(63,125,250,500,1000,2000,4000,8000),
    )+
  scale_color_manual(
    values = c("green", "blue", "orange"),
    breaks = c("0.5 p", "0.95 p", "0.99 p"))+
  labs(y = "P(|Binomial - Expected| >= |OOR - Expected|)",
       x = expression(paste("Frequency ",
                             italic(" f"), " / Hz")),
       color = "Predictive interval")+
  theme(legend.position = "bottom")


ppcFreqBinom

```



```{r}

oor <- c()

oor[1] <- OutOfRange(ppcDF, "NTP", "L1", "U1", "sum")
oor[2] <- OutOfRange(ppcDF, "NTP", "L2", "U2", "sum")
oor[3] <- OutOfRange(ppcDF, "NTP", "L3", "U3", "sum")


data.frame(
  "Expected" = alphas*21*142,
  "Out of range" = oor
)



```





## PPC Whiteining TVAR Mahalanobis 


```{r}

if (recalculate) {
  PPCwhiteSample <- apply(
      fit3seg$y_rep,
      1,
      whiteningTVAR,
      maxlag = 3
  )

  PPCwhiteSample <- t(PPCwhiteSample)
  saveRDS(PPCwhiteSample, file = "Saves/PPCwhite")
}



PPCwhiteSample <- readRDS("Saves/PPCwhite")



```


```{r}
ntp_datamat <- matrix(ntp_long$NTP, nrow = 142,byrow = T)

ntp_datawhite <- whiteningTVAR(ntp_datamat, 3)

```


```{r}


alphas <- c(0.5,0.05,0.01)

sim_quant <- apply(
  PPCwhiteSample, 2, quantile, probs = c(alphas/2, 1-alphas/2)
)

sim_quant <- data.frame(t(sim_quant))
names(sim_quant) <- c(
  paste("L", 1:length(alphas), sep = ""),
  paste("U", 1:length(alphas), sep = "")
)


own_ribbon <- data.frame(
  Ind = 1:142,
  ydata = ntp_datawhite,
  sim_mean = apply(PPCwhiteSample, 2, mean),
  sim_quant
)


basis <- ggplot(own_ribbon)+
    aes(x = Ind)+
    geom_ribbon(aes(ymin = L3, ymax = U3, fill = "99 p"), alpha = 0.1)+
    geom_ribbon(aes(ymin = L2, ymax = U2, fill = "95 p"), alpha = 0.2)+
    geom_ribbon(aes(ymin = L1, ymax = U1, fill = "50 p"), alpha = 0.3)+
    geom_line(aes(y = sim_mean, color = "PPC mean"))


PPCwhite <- basis +
    geom_point(aes(y = ydata, color = "Data"))+
    geom_line(aes(y= ydata, color = "Data"))+
    scale_fill_manual(values = rep("black", 3))+
    scale_color_manual(values = c("blue", "black"))+
    labs(color = "", fill = "Predictive interval",
         x = "Ceiling", y = "Squared mahalanobis")+
    theme(legend.position = "bottom")+
  scale_x_continuous(breaks = seq(10,142,10))

PPCwhite


```


```{r}


whichOOR3 <- !(own_ribbon$ydata >=own_ribbon$L3 & own_ribbon$ydata <=own_ribbon$U3)


inrangeWhite <- c()

inrangeWhite[1] <- sum(own_ribbon$ydata >=own_ribbon$L1 & own_ribbon$ydata <=own_ribbon$U1)
inrangeWhite[2] <- sum(own_ribbon$ydata >=own_ribbon$L2 & own_ribbon$ydata <=own_ribbon$U2)
inrangeWhite[3] <- sum(!whichOOR3)


data.frame(
  "Predictive interval" = alphas*142,
  "Out of range" = 142-inrangeWhite
)



```

```{r}

ntp_long$ID <- rep(1:142, each = 21)
ntp_long$mean <- rep(apply(lin_adjust, 1, mean), 142)


gghelp <- gg_basic(ntp_long[is.element(ntp_long$ID,which(whichOOR3) ),],
         aes(x = fpHz, y = NTP, group = ID),
         ymin = 0, ymax = 70, isl = T, lab = F)+
  geom_line(aes(color = as.factor(ID)),alpha = 0.5)+
  geom_point(aes(color = as.factor(ID)), size = 1/2)+
  geom_line(aes( y= mean, color = "Data mean"))+
  scale_color_manual(values = c(RavenR::ggplotColours(9), "black" ))+
  labs(color = "Ceiling")+
  theme(legend.position = "bottom")



mahalanobisOutlier1 <- gg_basic(ntp_long[is.element(ntp_long$ID,which(whichOOR3)[1:5]),],
         aes(x = fpHz, y = NTP, group = ID),
         ymin = 0, ymax = 70, isl = T, lab = F)+
  geom_line(aes(color = as.factor(ID)),alpha = 0.5)+
  geom_point(aes(color = as.factor(ID)), size = 1/2)+
  geom_line(aes( y= mean, color = "Data mean"))+
  scale_color_manual(values = c(RavenR::ggplotColours(9)[1:5], "black" ))+
  labs(color = "Ceiling")

mahalanobisOutlier2 <- gg_basic(ntp_long[is.element(ntp_long$ID,which(whichOOR3)[6:9]),],
         aes(x = fpHz, y = NTP, group = ID),
         ymin = 0, ymax = 70, isl = T, lab = F)+
  geom_line(aes(color = as.factor(ID)),alpha = 0.5)+
  geom_point(aes(color = as.factor(ID)), size = 1/2)+
  geom_line(aes( y= mean, color = "Data mean"))+
  scale_color_manual(values = c(RavenR::ggplotColours(9)[6:9], "black" ))+
  labs(color = "Ceiling")



ggarrange(
  mahalanobisOutlier1, mahalanobisOutlier2,
  legend = "bottom", legend.grob = get_legend(gghelp)
)


```





# Simulated example data

```{r}

createPublicData <- FALSE

if (createPublicData) {
  set.seed(26)


  public_data <- ntp_share
  public_data[public_data$seltsam_mbbm,] <- NA
  public_data[,-c(13:33)] <- NA
  
  
  y <- ntp_share[!ntp_share$seltsam_mbbm, 13:33]
  ycentered <- scale(y, scale = F)
  
  ymean <- apply(y, 2, mean)
  ycov <- resid_to_cov(ycentered, "tvAR", maxlag = 3)
  
  
  
  public_data[!ntp_share$seltsam_mbbm, 13:33] <- rmvnorm(
    142, ymean, ycov
  )
  
  
  write.csv2(public_data, "publicDataISO.csv", row.names = F)
}




```





